\section{The HARPO Agent}\label{app:arch}

HARPO normalizes a task bundle (\texttt{spec/constraints/budget} JSON)
into one \texttt{TaskContext}---top function, sources, testbench, part,
clock, policy, budget---all \emph{injected}, never hardcoded, so the same
agent runs an external evaluator's tasks unchanged. A tool runner
dispatches stages to a backend: host \texttt{g++} (functional csim, since
g++ ignores \texttt{\#pragma HLS}) or \texttt{vitis\_hls} (real
csim+csynth). A deterministic report parser maps logs to typed statuses; a
rule-based diagnosis engine maps a status to an action. Patch providers
share one protocol with three implementations: \texttt{recipe} (a
deterministic precise-pragma library, 0~tokens), \texttt{ollama} (a local
LLM), and \texttt{mock} (for tests). The two loops are \emph{repair} (csim
$\rightarrow$ diagnose $\rightarrow$ patch $\rightarrow$ re-csim) and
\emph{optimize} (baseline csim+csynth $\rightarrow$ propose pragma
$\rightarrow$ apply $\rightarrow$ re-verify csim $\rightarrow$ csynth
$\rightarrow$ keep iff the score strictly improves); \texttt{pipeline}
chains them on one shared budget.

In detail, HARPO is built from six cooperating components and two
control loops, chained by a pipeline that runs on a single shared per-task
budget account. This section details each component, the operational
steps of each loop, the \texttt{BudgetManager} that meters all tool
calls, the correctness-dominated lexicographic score, and the
correctness-preserving invariant that makes ``correctness before PPA''
structural rather than advisory.

\subsection{The six components}\label{app:arch:components}

\begin{enumerate}
  \item \textbf{Task loader.} Normalizes a task bundle
    (\texttt{spec.json} + \texttt{constraints.json} +
    \texttt{budget.json}) into one \texttt{TaskContext}: top function,
    source/testbench file lists, part, clock, policy, and budget. Part,
    clock, and sources are \emph{injected}, never hardcoded, so the same
    agent runs the evaluator's tasks unchanged.
  \item \textbf{Tool runner.} Dispatches a stage to a backend:
    \texttt{gpp} (host g++ compile+run = functional csim, no Vitis
    needed, since g++ ignores \texttt{\#pragma HLS}) or
    \texttt{vitis\_hls} (real csim+csynth via a generated
    \texttt{run\_harpo.tcl}).
  \item \textbf{Report parser.} \texttt{parse\_csim} $\rightarrow$
    \{pass, compile\_error, functional\_fail, timeout,
    tool\_unavailable\}; \texttt{parse\_csynth} $\rightarrow$ II, depth,
    latency best/worst, Fmax, LUT/FF/DSP/BRAM + util\% from the Vitis
    XML, with timing/resource violations.
  \item \textbf{Diagnosis engine.} Rule-based and deterministic (no
    model): maps a parsed status to a recommended action. A
    clean/improvable csynth pass yields \texttt{optimize\_ppa}; a
    repeated failure escalates to \texttt{rollback\_or\_escalate}.
  \item \textbf{Patch providers.} A uniform \texttt{PatchProvider}
    protocol with three implementations: \texttt{mock} (deterministic
    string edits, for tests/demo), \texttt{recipe} (deterministic
    precise-pragma library, no tokens), and \texttt{ollama} (local LLM
    over stdlib \texttt{urllib}, never raises---degrades to the next
    provider). Each reports \texttt{last\_usage} tokens.
  \item \textbf{Candidate store + lexicographic score.} Every attempt
    is an isolated, forkable copy of the source; the edited kernel is
    compiled against the \emph{un-edited} testbench; the winner is
    whichever candidate scores highest. The \texttt{BudgetManager} is
    the budget spine (per-tool limits + reserve + policy gates).
\end{enumerate}

\subsection{The two control loops}\label{app:arch:loops}

The agent runs two loops over the candidate store. The repair loop
restores correctness; the optimize loop improves power, performance, and
area (PPA) without ever breaking correctness. Rendered as step
sequences rather than a wide ASCII diagram, they are as follows.

\subsubsection{Repair loop (correctness)}
\begin{enumerate}
  \item Run csim via \texttt{gpp}, parse the report, and diagnose.
  \item If csim passes, stop: the current candidate is the winner
    (DONE).
  \item Otherwise check \texttt{budget.policy\_allows(llm)}.
  \item If allowed, call \texttt{provider.propose()} in provider order
    (\texttt{recipe} $\rightarrow$ \texttt{ollama}), accumulating
    tokens.
  \item Run \texttt{check\_contract} (signature / testbench / globals).
  \item If the contract holds, fork the source, apply the patch, and
    loop back to step~1.
  \item Stop on any of: csim pass; \texttt{max\_steps} reached; budget
    exhausted; or a repeated/regressed state.
\end{enumerate}

\subsubsection{Optimize loop (PPA, never breaks correctness)}
\begin{enumerate}
  \item Require the baseline csim to pass; otherwise return ``run repair
    first.''
  \item Run the baseline csynth to obtain \texttt{baseline\_metrics}.
  \item Loop while \texttt{steps < max} and
    \texttt{no\_improve < patience}:
  \begin{enumerate}
    \item \texttt{diagnose\_csynth} $\rightarrow$
      \texttt{provider.propose()}.
    \item \texttt{check\_contract}; fork; \texttt{apply\_patch}.
    \item \textbf{Re-verify csim} on the forked candidate; if it broke
      functional behavior, DISCARD the candidate.
    \item If csim still passes, run \texttt{csynth(child)}.
    \item Compute \texttt{score(child)}; if it is strictly greater than
      \texttt{score(best)}, ACCEPT, else REJECT.
  \end{enumerate}
\end{enumerate}

\subsubsection{Pipeline}
\texttt{run\_pipeline} chains repair then optimize on \textbf{one}
shared per-task \texttt{BudgetManager}, so repair and optimize debit the
same account.

\subsection{BudgetManager (the budget spine)}\label{app:arch:budget}

The \texttt{BudgetManager} is constructed from \texttt{budget.json}
(\texttt{mode: per\_tool}). Its per-tool limits are
\texttt{\{static\_check:100, csim:20, csynth:10, cosim:5,
llm\_calls:30\}}, with a held-back \texttt{reserve
\{final\_csim:1, final\_csynth:1, final\_cosim:1\}}. A missing limit is
unlimited. \texttt{can(action)} is true only while
$\texttt{spent} < \texttt{limit} - \texttt{reserve}$---the reserve
guarantees a winner can always be re-verified at the end.

\texttt{policy\_allows} enforces three gates, in order:
\begin{enumerate}
  \item \textbf{Budget} --- \texttt{can(action)} must hold before the
    call.
  \item \textbf{Stage ordering} --- no csynth/cosim before csim passes.
  \item \textbf{Stop/rollback guard} --- refuse another LLM call when
    the state is \emph{repeated} or \emph{regressed}.
\end{enumerate}

These ``don't-waste rules'' make tool calls---the scarce resource the
setting meters---the budgeted quantity: no csynth before csim
passes, and no fresh LLM call when re-trying a failing fix would burn
budget for nothing. The four demonstrated tasks of Sec.~\ref{app:suite} spent
\textbf{42 tool calls} and \textbf{2400 tokens} total---comfortably inside the
per-task limits (csim~20, csynth~10, llm\_calls~30).

\subsection{The correctness-dominated lexicographic score}\label{app:arch:score}

Ranking proceeds top-down so that correctness dominates PPA:
\begin{enumerate}
  \item \textbf{Correctness tier} --- 0: csim unknown/fail; 1: csim
    pass; 2: csim+csynth pass. A tier-2 candidate always outranks a
    tier-1, regardless of PPA.
  \item \textbf{Throughput on the design \texttt{interval\_max}} (within
    a tier) --- \emph{not} the per-loop II, which is diagnostic-only.
  \item \textbf{Objective-dependent secondary order.} The default
    objective is \texttt{satisfice\_then\_area}: once a candidate meets
    the per-task \texttt{throughput\_target} it is ranked by lowest
    normalized \texttt{area\_score}, then ADP (area $\times$ interval).
\end{enumerate}

The other objectives (\texttt{speed\_first}, \texttt{area\_first},
\texttt{adp}, \texttt{pareto\_report}) reorder these terms. The old
II-first, latency, LUT lexicographic order is no longer the live score;
ranking II ahead of latency is the right default for
streamed/throughput kernels but can \emph{raise single-call latency}
(on \texttt{matmul\_001}, \texttt{latency\_worst} rose
$260 \rightarrow 518$ even as II improved $4 \rightarrow 1$, because the
loop greedily takes the initiation-interval win), which is why the
per-task objective knob exists.

\subsection{The correctness-preserving invariant}\label{app:arch:invariant}

An optimization is accepted only if it (a)~still passes csim AND
(b)~strictly improves the score. Every forked candidate is re-run
through csim \emph{before} its csynth metrics are read; if a pragma
broke functional behavior, the candidate is discarded and its metrics
are never trusted. This is the structural guarantee behind
``correctness before PPA'': the agent can never trade a wrong-but-fast
design up the ranking. The invariant is proven by a hermetic test,
\texttt{tests/test\_optimize\_safety.py} with
\texttt{tasks/trap\_breakscsim\_001}; \textbf{106 unit tests} total, all
green.
