\section{Scoring Overhaul and Recipe-vs-LLM Analysis}\label{app:scoring}

This section develops the paper's central contribution: why a
deterministic recipe library beats a raw LLM on structured-reduction kernels, why
the area blow-up of the LLM arm was ultimately a \emph{scoring} defect rather than a
prompting defect, and how two distinct fixes (plus an autonomy probe) make the
elegant recipe outrank the LLM under honest scoring. All numbers are copied verbatim
from the committed evidence under \texttt{runs/} and \texttt{docs/ablations/};
the canonical 21-row evidence table is \texttt{docs/ablations/canonical/TABLE.md},
regenerable via \texttt{scripts/ablation\_table.py}.

\subsection{The recipe library}\label{app:scoring:recipes}

The deterministic \textbf{recipe library} (\texttt{recipes.py}) emits
\emph{fully specified}, correct-by-construction HLS pragmas, one per call, by robust
text scanning (\texttt{ARRAY\_PARTITION cyclic factor=N dim=1},
\texttt{PIPELINE II=1}, \texttt{UNROLL factor=N}, \ldots), each validated to compile by
an offline g++ syntax self-test. It costs \textbf{0 tokens}. The optimize default
provider order is \texttt{recipe,ollama} --- precise deterministic recipes first, the
LLM only for the tail the catalogue cannot reach (``dumbest reliable tool first'').

\subsection{The headline mac8 ablation}\label{app:scoring:headline}

Table~\ref{tab:headline} reports the headline ablation
(\texttt{docs/ablations/recipe-vs-llm.md}): the same optimize loop, the same kernel
\texttt{mac8\_001}, two providers, real Vitis HLS 2025.2.

\begin{table}[h]
\centering\small
\caption{Headline \texttt{mac8\_001} ablation: recipe vs.\ raw LLM, real Vitis HLS 2025.2.}
\label{tab:headline}
\begin{tabular}{lrrrrrr}
\toprule
design & II & latency & interval & LUT & FF & tokens \\
\midrule
baseline      & 4   & 1026 & 1024 & 369        & 153 & --- \\
recipe-best   & \textbf{1} & 259  & 256  & \textbf{315}   & 126 & \textbf{0} \\
ollama-best   & n/a$^{1}$ & \textbf{129} & 128  & \textbf{13194} & 322 & \textbf{34387} \\
\bottomrule
\end{tabular}

\vspace{2pt}
{\footnotesize $^{1}$ The LLM fully unrolls the outer loop, so Vitis reports no
per-iteration II (\texttt{ii: null}); throughput shows as total interval 128
(vs.\ baseline 1024) --- a real, correct throughput win, paid for in area. Both arms
re-verified csim-correct. The recipe-best move was a single
\texttt{ARRAY\_PARTITION cyclic factor=8 dim=1}; the LLM stacked
partition$\times$2 + \texttt{PIPELINE II=1} + \texttt{UNROLL factor=8}.}
\end{table}

This is \textbf{$\sim$42$\times$ area for comparable throughput} (recipe 315 LUT vs.\
LLM 13194 LUT; the LLM is in fact lower-latency, 129 vs.\ 259). For
\texttt{xc7z020} (53{,}200 LUT) the recipe sits at 0.6\% utilization, the LLM design
at 24\%. The LLM arm was run \textbf{3$\times$} and was \textbf{highly stable} ---
identical winning candidate, four-pragma stack, and PPA (LUT 13194, latency 129,
FF 322, 34{,}387 tokens) every time. The blow-up is consistent behavior, not a draw.

\subsection{The honest corrected cause: over-parallelization, enabled by the score}\label{app:scoring:cause}

A prior note framed the blow-up as an \emph{imprecise} \texttt{ARRAY\_PARTITION}
(missing the partition type) defaulting to \texttt{complete}. With the current
optimizer prompt (\texttt{\_OLLAMA\_OPT\_SYSTEM\_PROMPT}, which forbids bare
\texttt{ARRAY\_PARTITION} and warns that omitting the type ``detonates area'') the LLM
no longer makes that mistake --- every partition it emitted was a fully-specified
\texttt{cyclic factor=N}. The area blow-up \textbf{reproduced anyway, via a different
mechanism}: the LLM stacks \texttt{PIPELINE II=1} \emph{plus} \texttt{UNROLL factor=8}
on top of partitioning, fully spatially unrolling the reduction. Each step strictly
improves the lexicographic score (II $\rightarrow$ latency dominate area), so the
greedy loop accepts the ever-larger-but-faster design. The lesson, sharpened: a raw
LLM is fluent but imprecise about \emph{how hard to push} --- it over-applies
parallelism for marginal latency at order-of-magnitude area cost, while the recipe
applies exactly the one pragma that unblocks II=1 and stops.

Crucially, a follow-up re-ablation traced the \emph{enabler} of the blow-up to the
\textbf{score itself}, not the LLM. A fully-unrolled loop reports no per-loop II
(\texttt{None}), which the old throughput term sorted as neutral and thus ranked
\emph{above} a real II$\geq$1 --- so the loop literally rewarded over-unrolling (it
even accepted a design with interval 3073 vs.\ the baseline's 1024). Prompt hardening
alone does not fix this. Both durable fixes (\S\ref{app:scoring:fixes}) are now
implemented.

\subsection{The two fixes, in depth}\label{app:scoring:fixes}

The over-parallelization enabler is \textbf{two distinct defects} with \textbf{two
distinct fixes}, both implemented and re-baselined on real Vitis (106/106 unit tests
green).

\paragraph{Fix 1 --- metric: throughput is scored on the design \texttt{interval\_max}, not the per-loop \texttt{PipelineII}.}
The per-loop term sorted a fully-unrolled loop's missing II (\texttt{None}) as neutral
0, beating a real II$\geq$1 and rewarding over-unrolling. \texttt{interval\_max} is
always reported by Vitis; per-loop \texttt{ii} is now diagnostic-only. Under it, the
accepted \texttt{mac8\_001} interval-3073 design correctly loses to the baseline
(1024), and the old \texttt{conv2d\_001} over-pragma'd candidate (interval 328)
correctly loses to the baseline (191) --- while the proper recipe move now drives
conv2d to interval 82 under a target (\S\ref{app:scoring:rebaseline}).

\paragraph{Fix 2 --- policy: the default objective is \texttt{satisfice\_then\_area}.}
The per-task objective is a 5-value enum --- \texttt{speed\_first}, \texttt{area\_first},
\texttt{adp}, \texttt{satisfice\_then\_area} (default), \texttt{pareto\_report}
(\texttt{spec.json} \texttt{objective}; legacy \texttt{throughput}/\texttt{latency}
alias to \texttt{speed\_first}; unknown/absent $\rightarrow$ default).
\texttt{interval\_max} scoring alone does not stop a \emph{genuinely faster but huge}
design from winning on throughput; satisficing throughput to a per-task
\texttt{throughput\_target} and then minimizing area is what makes the elegant recipe
(interval 256 / 315 LUT) outrank the LLM blow-up (128 / 13194 LUT).

Supporting these, a new \texttt{harpo/area.py} provides a normalized
\texttt{area\_score} (used/available summed over LUT/FF/DSP/BRAM, with \emph{no
per-resource weights} so scarcity is not double-counted; per-part fallback for
xc7z020), \texttt{adp = area\_score $\times$ interval\_max}, \texttt{resource\_growth\_ratio},
and a \texttt{pareto\_front} helper for the reporting mode. A new per-task
\texttt{throughput\_target} field carries the \texttt{interval\_max} ceiling the
satisfice policy targets.

\paragraph{Fix 3 --- autonomy: a recipe-only capped probe derives the target when the task gives none.}
\texttt{satisfice\_then\_area} needs a \texttt{throughput\_target}; rather than require
a hand-written one, \texttt{harpo/probe.py} derives it. The probe is deliberately
minimal: \textbf{recipe-only, zero LLM tokens}, capped at \texttt{max\_synth}
single-pragma forks from the baseline (default 4), \textbf{full unroll excluded}. The
target is the lowest \texttt{interval\_max} among the baseline and any probe candidate
that csim+csynth-passes \emph{and} stays within \texttt{area\_score} $\leq$ 2$\times$
baseline; if nothing beats the baseline within that cap, it falls back to the baseline
\texttt{interval\_max} (never an unreachable goal). On \texttt{mac8\_001} with its
target stripped, the probe derives \textbf{256} --- exactly the hand-set value --- at
0 tokens.

\subsection{Re-baseline under the new scoring}\label{app:scoring:rebaseline}

Recipe arm, real Vitis 2025.2, \texttt{xc7z020clg400-1}, 10 ns. Every number here is a
row in the canonical evidence table (\texttt{docs/ablations/canonical/TABLE.md}). The
two fixes show on two kernels.

\paragraph{\texttt{mac8} (Fix 1, metric).} Full unroll makes \texttt{interval\_max}
\emph{worse} (3073) $\rightarrow$ Fix 1 discards it, and the recipe reaches interval
256 / LUT 315.

\paragraph{\texttt{matmul} (Fix 2, policy).} Full unroll makes \texttt{interval\_max}
genuinely \emph{better}, so the metric fix alone will not stop it. The two matmul
recipe rows make the case directly (Table~\ref{tab:matmul}):
\texttt{speed\_first} reaches interval 19 / LUT 5689 / FF 14999 (area\_score 1.121,
ADP 21.292), while \texttt{satisfice\_then\_area} (target 72) keeps interval 44 /
LUT 3121 / FF 5932 (area\_score 0.3326, ADP 14.6344) --- $\sim$45\% fewer LUT,
$\sim$60\% fewer FF, \emph{and} a lower ADP, for 25 cycles of interval. The target is
what governs the over-push.

\begin{table}[h]
\centering\small
\caption{\texttt{matmul\_001} recipe arm: \texttt{satisfice\_then\_area} (target 72) vs.\ \texttt{speed\_first}.}
\label{tab:matmul}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
policy & interval & LUT & FF & area\_score & ADP \\
\midrule
baseline                & 256 & 706  & 579   & 0.04599 & 11.7722 \\
satisfice (target 72)   & 44  & \textbf{3121} & \textbf{5932}  & 0.3326  & \textbf{14.6344} \\
speed\_first            & \textbf{19} & 5689 & 14999 & 1.121   & 21.292 \\
\bottomrule
\end{tabular}}
\end{table}

\paragraph{PolyBench (Fix 3, probe).} Three PolyBench kernels (gemm/atax/bicg, integer
16$\times$16) extend the evidence beyond hand-built toys; all three auto-derive their
target via the Fix-3 probe (Table~\ref{tab:polybench}). \texttt{gemm} conservatively
falls back to baseline 2060 (no improvement, baseline kept); \texttt{atax} derives a
real target of 280 and reaches interval 64 / LUT 3907; \texttt{bicg} derives 162 and
reaches interval 60 / LUT 3596 --- both optimizing within a 2$\times$ area cap.

\begin{table}[h]
\centering\small
\caption{PolyBench recipe arm (16$\times$16 integer), targets auto-derived by the Fix-3 probe.}
\label{tab:polybench}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llrrrr}
\toprule
kernel & target & interval & LUT & FF & ADP \\
\midrule
gemm\_001 & 2060 (fallback)     & 2060 & 1238 & 1180 & 126.965 \\
atax\_001 & 280 (auto-derived)  & 64   & 3907 & 5596 & 22.0298 \\
bicg\_001 & 162 (auto-derived)  & 60   & 3596 & 8929 & 35.2726 \\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Recipe-vs-LLM comparison under honest scoring}\label{app:scoring:comparison}

Table~\ref{tab:recipe-vs-llm} is the head-to-head, recipe vs.\ raw LLM under the
corrected scoring (the four re-baselined LLM arms). On structured-reduction kernels the
raw LLM over-parallelizes and, under the corrected scoring, \textbf{wins nothing}; the
precise recipe applies the single unblocking pragma. On kernels the safe recipes cannot
crack both are null; elsewhere they are comparable. The \texttt{recipe}-before-\texttt{ollama}
default is what captures the best of both.

\begin{table*}[t]
\centering\small
\caption{Recipe vs.\ raw LLM under honest (\texttt{interval\_max} + \texttt{satisfice\_then\_area}) scoring.}
\label{tab:recipe-vs-llm}
\begin{tabular}{llll}
\toprule
kernel & recipe arm & raw-LLM arm & winner \\
\midrule
mac8\_001   & interval 256 / LUT 315                          & no improvement (full-unroll interval 3073 discarded) & \textbf{recipe} \\
matmul\_001 & interval 44 / LUT 3121 \emph{(satisfice, target 72)} & no improvement (no candidate meets target)          & \textbf{recipe} \\
gemm\_001   & no improvement (probe fallback, target 2060)    & no improvement                                       & tie (both null) \\
atax\_001   & interval 64 / LUT 3907 / ADP 22.0               & interval 81 / LUT 3994 / ADP 28.0                    & recipe (marginal) \\
\bottomrule
\end{tabular}
\end{table*}

LLM-arm re-baselines for the remaining hand-built kernels, and additional PolyBench
ports, remain future work; the released evidence covers the four head-to-head kernels
above.
