\section{Reproducibility}\label{app:repro}

Everything in this paper runs on \emph{free, local} infrastructure at \$0 LLM
cost, and every reported number is regenerable from committed, replayable JSON
artifacts in the public repository.

\subsection{Environment}

The LLM is a local \textbf{qwen} model served by Ollama over plain HTTP; the
endpoint and model are read from the environment variables
\texttt{HARPO\_OLLAMA\_URL} and \texttt{HARPO\_OLLAMA\_MODEL}, so there
is no per-token cost. Synthesis uses \textbf{Vitis HLS 2025.2}, which is free on
Linux for releases prior to 2026.1. The target part is
\texttt{xc7z020clg400-1} (an entry-level device covered by the free tier) at a
10\,ns clock. All experiments were run on a Linux workstation. The agent package
itself is \textbf{stdlib-only} (no third-party Python dependencies).

\subsection{Command one-liners}

Source the Vitis environment once before any csynth/optimize/pipeline command;
the repair, optimize, and pipeline entry points then drive the full flow, and
\texttt{run\_suite.py} aggregates per-run results into a summary table.

{\footnotesize
\begin{verbatim}
# once, before csynth/optimize/pipeline:
source ~/tools/Xilinx/2025.2/Vitis/settings64.sh

python3 -m harpo repair \
  tasks/vadd_buggy_001 --provider mock,ollama
python3 -m harpo optimize \
  tasks/mac8_001 --provider recipe,ollama
# repair, then optimize:
python3 -m harpo pipeline \
  tasks/vadd_buggy_001
# aggregate runs/ -> SUITE.md + SUITE.csv:
python3 scripts/run_suite.py
\end{verbatim}}

The headline ablation compares the recipe arm against the raw-LLM arm on the
same kernel; each invocation writes its result JSON under
\texttt{docs/ablations/}:

{\footnotesize
\begin{verbatim}
# the headline ablation
# (each writes docs/ablations/mac8_001_*.json):
python3 -m harpo optimize \
  tasks/mac8_001 --provider recipe
python3 -m harpo optimize \
  tasks/mac8_001 --provider ollama
\end{verbatim}}

The offline self-tests require neither Vitis nor the LLM; they validate the
csim classifier, the recipe provider's emitted C++, and the csynth parser
against stored artifacts:

{\footnotesize
\begin{verbatim}
# offline self-tests (no Vitis, no LLM):
# csim classification / recipe C++ validity
# (g++ -fsyntax-only) / csynth XML parsing
python3 scripts/selftest.py
python3 scripts/selftest_recipes.py
python3 scripts/selftest_csynth.py
\end{verbatim}}

\subsection{Replayable per-run artifacts}

Every run writes replayable JSON under \texttt{runs/<task\_id>/}: for each
candidate, the \texttt{\{csim,csynth\}\_\{raw,parsed\}.json} files, plus the
phase log carrying the budget, token counts, and per-candidate scores. Because
each candidate's raw and parsed synthesis output is committed alongside the code,
every number in this paper can be re-derived from these logs.

\subsection{Install caveat}

The unified Vitis 2025.2 installer omits the \texttt{bin/vitis\_hls} launcher.
The loader wrapper must be recreated by hand; otherwise \texttt{vitis\_hls}
is not on \texttt{PATH} even after sourcing \texttt{settings64.sh}.
